  
---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
# getarc
  <!-- badges: start -->
  [![Codecov test coverage](https://codecov.io/gh/MatthewJWhittle/getarc/branch/master/graph/badge.svg)](https://codecov.io/gh/MatthewJWhittle/getarc?branch=master)
  
  <!-- badges: end -->

# Overview
`getarc` is an R wrapper for the [ArcGIS Rest API](https://developers.arcgis.com/rest/services-reference/). It provides access to the extensive open data available from [Arc GIS](https://hub.arcgis.com/search).
It currently only supports functionality for querying data. 

* `query_layer` gets data from an arc gis server and supports query operations
* `get_layer_details` gets metadata about a layer such as the field names and maxRecordCount
* `get_token` gets an access token via a web browser login to access private services

# Installation
The package can currently be installed from github:
```{r eval=FALSE}
# Install the development version from GitHub:
install.packages("devtools")
devtools::install_github("matthewjwhittle/getarc")
```
```{r include=FALSE}
devtools::load_all()
```

# Examples
```{r warning=FALSE, message=FALSE}
library(getarc)
library(sf)
library(tidyverse)
```

# Getting data
Data on an arc gis server can be accessed via the `query_layer` function. This function also supports any query operation supported by the ArcGIS Rest API.

# Getting data from a Feature Server - National Parks in England
```{r}
# Use the endpoints list 

# Get the data
national_parks <- 
  query_layer(endpoint = endpoints$national_parks_england)

head(national_parks)
# Plot the first feature
plot(national_parks$geometry[1])
```

## Querys
A query can be included in the request by either supplying a named list to the `query` parameter, or passing a bounding box to `bounding_box` to return intersecting features.

The query parameter supports any query parameter supported by the API Please review the [API documentation](https://developers.arcgis.com/rest/services-reference/query-feature-service-layer-.htm) on querying Feature Server layers for detail on how to query data. I intend to provide more R-friendly support for query operations in the future.

Returning only one feature.
```{r}
one_park <- 
  query_layer(endpoint = endpoints$national_parks_england,
  # Return only one record
  query = c(resultRecordCount = 1)
)
print(one_park)
plot(one_park$geometry)

# Including a sql where query to only return the yorkshire dales
yorkshire_dales <- 
  query_layer(endpoint = endpoints$national_parks_england,
  # SQL query to return data for the yorkshire dales 
  query = c("where" = "NAME LIKE 'YORKSHIRE DALES'")
)
```

Spatial querys can be perform either using the `in_geometry` and `spatial_filter` arguments. These arguments define a geometry spatial filter (intersects, contains, etc.) to query the endpoint. It is possible to pass a bounding box or polygon to `in_geometry`. Complex polygons are simplified to reduce the number of characters being sent in the request.

```{r}
dales_bbox <- st_bbox(yorkshire_dales)

# Which Sites of Special Scientific Interest are in the yorkshire dales?
dales_sssi <- 
  query_layer(endpoints$sssi_england, 
  # Supply a polygon for a spatial intersects query
  in_geometry = yorkshire_dales
)

# Transform the data for plotting
yorkshire_dales <- yorkshire_dales %>% st_transform(crs = 27700)
dales_sssi <- dales_sssi %>% st_transform(crs = 27700)

# Plot the yorkshire dales and it's SSSIs
plot(yorkshire_dales$geometry)
plot(dales_sssi$geometry, add = TRUE, border = "red", col = "orange")
```

# Endpoints 
The endpoint url can be copied from Query URL box of th Arc GIS API explorer. A number of endpoints are provided in `getarc::endpoints`.
```{r}
head(endpoints)
```

# Don't Return Geometry
Often it is desirable to avoid returning the geometry and download a table of results to investigate. This can then be used to build up queries to refine your results, improving download times. This is done using `return_geometry = FALSE`
```{r}
query_layer(endpoint = endpoints$national_parks_england, 
            return_geometry = FALSE)
```

# Custom Queries
A number of query options are supported by esri that I haven't yet built an R api for. You can take advantage of this by passing a named vector of query parameters and values to the `query` argument. 

SQL WHERE queries are supported via this argument:
```{r}
query_layer(endpoints$gb_postcodes, 
            query = c(resultRecordCount = 1))

south_harrogate_postcodes <- 
  query_layer(endpoints$gb_postcodes, 
            query = c(where = "pcd LIKE 'HG2%'"))

# Plot the postcodes
south_harrogate_postcodes %>% 
  # Extract the postcode sector (HG1 1)
  mutate(postcode_sector = str_extract(pcd, "[A-Z]{1,2}[0-9]+ [0-9]+")) %>% 
  select(postcode_sector) %>% plot()
```

# Authentication
Authentication is done via oauth2.0 and the `get_token` function. Sign in is done via a popup browser window and the token can be passed to the query functions via the `my_token` argument. To use authentication you need to [set up an arcgis developers account](https://developers.arcgis.com/sign-up/) (for free) and create an app. This takes about 5 minutes. Once complete you need to navigate to the app on your dashboard, copy the credentials and set them using `set_credentials`. These are saved securely as environment variables and now you can use `get_token` to access secured services where you have access. 
```{r, eval=FALSE}
set_credentials(client_id = "xxxx", client_secret = "xxxx", app_name = "My App")

my_token <- get_token()

data <-
  query_layer(endpoint = private_endpoint,
  # Pass in token for a secured service
  my_token = my_token
)
```



