---
title: "Making-Queries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Making-Queries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(getarc)
library(devtools)
load_all()
library(sf)
library(tidyverse)

```

Data on an arc gis server can be accessed via the `query_layer` function. This function also supports any query operation supported by the ArcGIS Rest API.

# Getting data from a Feature Server - National Parks in England

## Querys

A query can be included in the request by either supplying a named list to the `query` parameter, or passing a bounding box to `bounding_box` to return intersecting features.

The query parameter supports any query parameter supported by the API Please review the [API documentation](https://developers.arcgis.com/rest/services-reference/query-feature-service-layer-.htm) on querying Feature Server layers for detail on how to query data. I intend to provide more R-friendly support for query operations in the future.

Returning only one feature.

```{r}

one_park <- 
  query_layer(endpoint = endpoints$national_parks_england,
  # Return only one record
  query = c(resultRecordCount = 1)
)
print(one_park)
plot(one_park$geometry)

# Including a sql where query to only return the yorkshire dales
yorkshire_dales <- 
  query_layer(endpoint = endpoints$national_parks_england,
  # SQL query to return data for the yorkshire dales 
  query = c("where" = "NAME LIKE 'YORKSHIRE DALES'")
)
```

Spatial querys can be perform either using the `in_geometry` and `spatial_filter` arguments. These arguments define a geometry spatial filter (intersects, contains, etc.) to query the endpoint. It is possible to pass a bounding box or polygon to `in_geometry`. Complex polygons are simplified to reduce the number of characters being sent in the request.

```{r}
dales_bbox <- st_bbox(yorkshire_dales)

# Which Sites of Special Scientific Interest are in the yorkshire dales?
dales_sssi <- 
  query_layer(endpoints$sssi_england, 
  # Supply a polygon for a spatial intersects query
  in_geometry = yorkshire_dales
)

# Transform the data for plotting
yorkshire_dales <- yorkshire_dales %>% st_transform(crs = 27700)
dales_sssi <- dales_sssi %>% st_transform(crs = 27700)

# Plot the yorkshire dales and it's SSSIs
plot(yorkshire_dales$geometry)
plot(dales_sssi$geometry, add = TRUE, border = "red", col = "orange")
```

# Endpoints

The endpoint url can be copied from Query URL box of th Arc GIS API explorer. A number of endpoints are provided in `getarc::endpoints`.

```{r}
head(endpoints)
```

# Don't Return Geometry

Often it is desirable to avoid returning the geometry and download a table of results to investigate. This can then be used to build up queries to refine your results, improving download times. This is done using `return_geometry = FALSE`

```{r}
query_layer(endpoint = endpoints$national_parks_england, 
            return_geometry = FALSE)
```

# Custom Queries

A number of query options are supported by esri that I haven't yet built an R api for. You can take advantage of this by passing a named vector of query parameters and values to the `query` argument.

SQL WHERE queries are supported via this argument:

```{r}
query_layer(endpoints$gb_postcodes, 
            query = c(resultRecordCount = 1))

south_harrogate_postcodes <- 
  query_layer(endpoints$gb_postcodes, 
            query = c(where = "pcd LIKE 'HG2%'"))

# Plot the postcodes
south_harrogate_postcodes %>% 
  # Extract the postcode sector (HG1 1)
  mutate(postcode_sector = str_extract(pcd, "[A-Z]{1,2}[0-9]+ [0-9]+")) %>% 
  select(postcode_sector) %>% plot()
```

# 
